---
title: "FinalProjectForecast"
author: "Duc Nguyen"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


#1 Project Setup
```{r}
url <- "https://raw.githubusercontent.com/nguye2dc/Inflation-Forecast-final/main/US%20CPI.csv"
```


```{r}
t <- read.csv(url, header =TRUE)
```
```{r}
head(t)
```
#2 Introduction
Data set is U.S. consumer price index (inflation) from 1913-2021.I found it on Kaggle. The dataset has CPI as a forecast variable and Date(Yearmon) as an index. The data set has one time series data on CPI and it would be interesting to see the prediction of CPI in the future.The forecast on this dataset can be used to predict the raise in Consumer Price Index nationally which would lead to understanding of where the economy is going.


#3 Data wrangling
```{r}
t$Yearmon = as.Date(t$Yearmon,"%m-%d-%Y")
```

```{r}
head(t)
```
```{r}
library(dplyr)
library(tidyverse)
library(lubridate)
library(tsibble)
```



```{r}
newDF <-t %>% as_tsibble(index = Yearmon)
```
```{r}
newDF2<-newDF %>% index_by(year=year(Yearmon)) %>% summarise(aveyearlyCPI= mean(CPI))
```

```{r}
index(newDF2)
```
```{r}
library(fpp3)
```

#4 Exloratpry analysis of the data set
```{r}
newDF2 %>% autoplot()
```
Seems like there are no seasonality with a huge linear trend for CPI
```{r}
library(ggplot2)
library(fpp3)
```

#5 Model fitting
```{r}
train <- newDF2 %>% filter(year<=1995)
test <- newDF2 %>% filter(year>1995)
```
train set is 80% of the data and 20% is test set.

```{r}
fit <- train %>% 
  model(
  autoLM = TSLM(aveyearlyCPI),
  autoETS = ETS(aveyearlyCPI),
  autoArima = ARIMA(aveyearlyCPI),
  LMwithTrend = TSLM(aveyearlyCPI ~ trend()),
  mean1= MEAN(aveyearlyCPI),
  naive= NAIVE(aveyearlyCPI),
  Arima = ARIMA(aveyearlyCPI~ pdq(0,2,2)),
  hwets = ETS(aveyearlyCPI~ error("A")+ trend("A")+ season("N"))
)
```

```{r}
glance(fit)
```
ARIMA model families has the lowest AICc
Only autoLM and LMwithTrend models have residuals

#Accuracy and forecast
```{r}
fc <- fit %>% forecast(h="30 years") 
fc %>% autoplot(newDF2,level = NULL)
```
forecast 30 years into the future because that is long enough to see how well the forecast fit into the whole data set using only the training data . 
Something to watch out for when putting this forecast into production is that there could be an unexpected circumstance that happens such as an epidemic or a war that could affect the CPI of the whole country.
```{r}
fc %>% accuracy(newDF2)
```

lower MAPE is better. Same for MAE, MPE because they compare the accuracy among different models

I choose the ARIMA model as the final model to forecast because it has the lowest MAPE, as well as MPE









